---
title: "ica-wrangling"
format: html
---

```{r}
library(tidyverse)
elections <- read.csv("https://mac-stat.github.io/data/election_2020_county.csv")
```

## Exercise 1: select Practice
```{r}
# Define elections_small
elections_small <- elections %>% 
  select(state_name, county_name, total_votes_20, repub_pct_20, dem_pct_20, total_votes_16, dem_pct_16)

# Check out the first 6 rows to confirm your code did what you think it did!
head(elections_small)
```

## Exercise 2: filter Demo

```{r}
# Keep only data on counties in Hawaii
elections_small |>
 filter(state_name == "Hawaii")
```

```{r}
# What does this do?
elections_small |>
  filter(state_name %in% c("Hawaii", "Delaware"))
```
```{r}
# Keep only data on counties where the Republican got MORE THAN 93.97% of the vote in 2020
# THINK: What variable is relevant here?
elections_small |>
  filter(repub_pct_20 >= 93.97)
```

```{r}
# Keep only data on counties in Texas where the Democrat got more than 65% of the vote in 2020
# Do this 2 ways.
# Method 1: 2 filters with 1 condition each
elections_small |>
 filter(state_name == "Texas") |>
 filter(dem_pct_20 > 65)

# Method 2: 1 filter with 2 conditions
elections_small |>
 filter(state_name == "Texas", dem_pct_20 > 65)
```

## Exercise 3: arrange Demo

```{r}
# Arrange the counties in elections_small from lowest to highest percentage of 2020 Republican support
# Print out just the first 6 rows
elections_small |>
  arrange(repub_pct_20) |>
  head()
```

```{r}
# Arrange the counties in elections_small from highest to lowest percentage of 2020 Republican support
# Print out just the first 6 rows

elections_small |>
  arrange(desc(repub_pct_20)) |>
  head()
```

## Exercise 4: mutate Demo
```{r}
# What did this code do?
elections_small |>
  mutate(diff_20 = repub_pct_20 - dem_pct_20) |>
  head()
```
Creates a new column with the difference between democrat and republican vote %

```{r}
# What did this code do?
elections_small |>
  mutate(repub_votes_20 = round(total_votes_20 * repub_pct_20/100)) |>
  head()
```
Creates a new column with the actual number of republican votes not just a percentage

```{r}
# What did this code do?
elections_small |>
  mutate(repub_win_20 = repub_pct_20 > dem_pct_20) |>
  head()
```
Creates a new column that shows weather binary indicator of weather the republicans won the county

```{r}
# You try
# Define a variable that calculates the change in Dem support in 2020 vs 2016
elections_small |>
 mutate(dem_change = dem_pct_20 - dem_pct_16) |>
 head()
```

```{r}
# You try
# Define a variable that determines whether the Dem support was higher in 2020 than in 2016 (TRUE/FALSE)
elections_small |>
 mutate(dem_improve = dem_pct_20 > dem_pct_16) |>
 head()
```

## Exercise 5: Pipe Series
# Part a
```{r}
elections_small |>
  filter(state_name == "Wisconsin",
         repub_pct_20 < dem_pct_20) |>
  arrange(desc(total_votes_20)) |>
  head()
```
# Part b
```{r}
# Now try it. Change the order of filter and arrange below.
elections_small |>
  arrange(desc(total_votes_20)) |>
  filter(state_name == "Wisconsin",
         repub_pct_20 < dem_pct_20) |>
  head()
```

# Part c
So the order of filter() and arrange() did not matter – rerranging them produces the same results. BUT what is one advantage of filtering before arranging?

- an advantage of filtering before aranging is there are less rows to arrange so its an easier function?

# Part d
```{r}
elections_small |>
  filter(state_name == "Delaware") |>
  mutate(repub_win_20 = repub_pct_20 > dem_pct_20) |>
  select(county_name, repub_pct_20, dem_pct_20, repub_win_20)
```

# Part e
```{r}
# Now try it. Change the order of mutate and select below.
# elections_small |>
#   filter(state_name == "Delaware") |>
#   select(county_name, repub_pct_20, dem_pct_20, repub_win_20)
  # mutate(repub_win_20 = repub_pct_20 > dem_pct_20) |>
```
There is an error beacuse we try and select for a variable that we dont have. 

## Exercise 6: DIY Pipe Series

# Part a
```{r}
elections_small |>
  filter(state_name == "Minnesota") |>
  arrange(desc(dem_pct_20)) %>% 
  select(county_name, dem_pct_20)
```
# Part b
```{r}
mn_wi <- elections_small %>% 
  filter(state_name %in% c("Minnesota", "Wisconsin")) %>% 
  mutate(dem_change = dem_pct_20 - dem_pct_16) %>% 
  arrange(dem_change) %>% 
  select(state_name, county_name, dem_pct_20, dem_pct_16, dem_change)

head(mn_wi)
  
```
# Part c
```{r}
ggplot(mn_wi, aes(x = dem_change, fill = state_name)) +
  geom_density(alpha = 0.7)
```
## Exercise 7: summarize Demo
```{r}
# What does this do?
elections_small |>
  summarize(median(repub_pct_20))
```
finds the median republican percentage for all US counties in 2020

```{r}
# What does this do?
elections_small |>
  summarize(median_repub = median(repub_pct_20))
```
makes the title more intelegable

```{r}
# What does this do?
elections_small |>
  summarize(median_repub = median(repub_pct_20), total_votes = sum(total_votes_20))
```
it adds the summary of the total votes in the us in 2020

## Exercise 8: summarize + group_by demo

```{r}
elections_small |>
  group_by(state_name)
```

```{r}
# Check out the structure before and after group_by
elections_small |>
  class()

elections_small |>
  group_by(state_name) |>
  class()
```

