---
title: "ica-dates"
format: html
---

## Exercise 1: More Filtering

```{r}
library(tidyverse)
penguins <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-28/penguins.csv')

# Check it out
head(penguins)
```


```{r}
# Create a dataset with just Adelie and Chinstrap using %in%
# Pipe this into `count(species)` to confirm that you only have these 2 species
penguins |>
  filter(species %in% c("Adelie", "Chinstrap")) |>
  count(species)
```
```{r}
penguins |>
  filter(species != "Gentoo") |>
  count(species)
```
```{r}
# NOTE the use of is.na()
penguins_w_body_mass <- penguins |> 
  filter(!is.na(body_mass_g))

# Compare the number of penguins in this vs the original data
nrow(penguins_w_body_mass)
```

We should only use na.omit in extreme circumstances beacuse if there are NAs in a column we are not focusing on, it will still get rid of them even though it wasnt necessary.

## Exercise 2: More Selecting

```{r}
# First: recall the variable names
names(penguins)
```

```{r}
# Use a shortcut to keep everything but the year and island variables
penguins %>% 
  select(-island, -year)
```

```{r}
# Use a shortcut to keep only species and the penguin characteristics measured in mm
penguins %>% 
  select(contains("mm"))
```

```{r}
# Use a shortcut to keep only species and bill-related measurements
penguins %>%
  select(species, contains("bill"))
```
```{r}
# Use a shortcut to keep only species and the length-related characteristics
penguins %>%
  select(species, contains("length"))
```

## Exercise 3: Arranging, Counting, & Grouping by Multiple Variables
```{r}
# Change this code to sort the penguins by species, and then island name
# NOTE: The first row should be an Adelie penguin living on Biscoe island
penguins |> 
  arrange(species, island)
```
```{r}
# Change this code to count the number of male/female penguins observed for each species
penguins |> 
  count(species, sex)
```

```{r}
# Change this code to calculate the average body mass by species and sex
penguins |> 
  group_by(species, sex) |> 
  summarize(mean = mean(body_mass_g, na.rm = TRUE))
```

## Exercise 4: Dates
```{r}
# Get today's date
as.Date(today())
```

```{r}
# Let's store this as "today" so we can work with it below
today <- as.Date(today())

# Check out the class of this object
class(today)
```

```{r}
year(today)
```

```{r}
# What do these lines produce / what's their difference?
month(today)
```

```{r}
month(today, label = TRUE)
```

```{r}
# What does this number mean?
week(today)
```

it means we are 9 weeks into the year

```{r}
# What do these lines produce / what's their difference?
mday(today)
```
26 days into the month

```{r}
yday(today)  # This is often called the "Julian day"
```

57 days into the year

```{r}
# What do these lines produce / what's their difference?
wday(today)
```

4 days into the week

```{r}
wday(today, label = TRUE)
```

```{r}
# What do the results of these 2 lines tell us?
today >= ymd("2024-02-14")
```

```{r}
today < ymd("2024-02-14")
```

## Exercises Part 2: Application
```{r}
library(mosaic)
data("Birthdays")
head(Birthdays)
```

## Exercise 5: Warming up
```{r}
# How many days of data do we have for each state?
Birthdays %>% 
  count(state)

# How many total births were there in this time period?
Birthdays %>% 
  summarise(total_births = sum(births))

# How many total births were there per state in this time period, sorted from low to high?
Birthdays %>% 
  group_by(state) %>% 
  summarise(total_births = sum(births)) %>% 
  arrange(desc(total_births))
```
## Exercise 6: Homework Reprise
```{r}
daily_births <- Birthdays %>% 
  group_by(date) %>% 
  mutate(wday = wday(date, label = TRUE)) %>% 
  summarise(total_births = sum(births))
head(daily_births)
```

